robustness-check

### 2025-04-19 code/analysis-code/statistical-analysis.qmd
project #5

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(broom)      
library(kableExtra) 
library(webshot)
library(magick)
library(here)
library(modelsummary)
library(plm)
library(lmtest)
library(gt)
library(webshot2)
library(fixest)
library(lubridate)
library(Metrics)
library(kableExtra)
library(here)

@@

@@

Grouping the variables.

@@

@@

```{r}
# Save the table as a HTML file
html_file <- here("results", "tables", "regression_results.html")
save_kable(table_html, file = html_file)

@@

@@

@@

@@

@@

# creating and save HTML table
unit_root_table %>%
  kbl(format = "html", digits = 4, align = "c",
      caption = "Primary model Unit Root Test Results") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  add_footnote("Null hypothesis: All series contain unit roots", 
              notation = "none") %>%
  save_kable(here("results", "tables", "unit_root_results.html"))
```

@@

>>>>

```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(broom)      
library(webshot)
library(magick)
library(here)
library(modelsummary)
library(plm)
library(lmtest)
library(gt)
library(webshot2)
library(fixest)
library(lubridate)
library(Metrics)
library(kableExtra)
library(here)

@@

# Importing datasets

@@

# Grouping the variables.

@@

# Preliminary Regression

@@

```{r}
# Save the table as a HTML file
html_file <- here("products", "supplementary-files", "regression_results.html")
save_kable(table_html, file = html_file)

@@

# Suggested Model

@@

# Primary regression model

@@

# Null Model

@@

# ADF test

@@

# creating and save HTML table
unit_root_table %>%
  kbl(format = "html", digits = 4, align = "c",
      caption = "") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  add_footnote("", 
              notation = "none") %>%
  save_kable(here("results", "tables", "unit_root_results.html"))
```

@@

# Regression model without Life Expectancy

### 2025-04-19 code/eda-code/eda.qmd
project #5

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(patchwork)
library(kableExtra)
library(htmltools)
library(magrittr)
library(webshot2)
library(patchwork)
```

@@

# Create a summary table, excluding specific rows
country_summary_table <- data_2010 %>%
  filter(!CountryName %in% c("Data from database: World Development Indicators", "OECD members",        "Last Updated: 01/28/2025")) %>%
  select(
    CountryName,
    GDP_current_US,
    Population_total,
    Life_expectancy_at_birth_total_years,
    Physicians_per_1000_people,
    Hospital_beds_per_1000_people,
    Government_Effectiveness_Estimate,
    Control_of_Corruption_Estimate,
    Rule_of_Law_Estimate,
    Access_to_electricity__of_population,
    Unemployment_total__of_total_labor_force_modeled_ILO_estimate,
    Poverty_headcount_ratio_at_national_poverty_lines__of_population
  )

# Generate the HTML table
table_html <- country_summary_table %>%
  kable("html", escape = FALSE, digits = 2) %>%
  kable_styling(full_width = FALSE, position = "left") %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c(" " = 1, "National Development Indicators" = 11)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Save the table object as an RDS file
rds_file <- here::here("results", "tables", "health_exp_2010.rds")
saveRDS(table_html, file = rds_file)

@@

# Creating the line plot
ggplot(data_filtered, aes(x = Year, y = Current_health_expenditure__of_GDP, color = CountryName, group = CountryName)) +
  geom_line(size = 1) +
  geom_point(size = 2) +  
  labs(
    title = "Current Health Expenditure as a % of GDP (2010–2020): Top 5 nations",
    x = "Year",
    y = "Health Expenditure (% of GDP)",
    color = "Country"
  ) +
  scale_y_continuous(breaks = y_breaks) +  
  theme_minimal()

@@

The Health Expenditure Per Capita (USD) varriable in fact shows more variation and also show some connections with annual population growth.
```{r}

@@

# Creating the line plot
ggplot(data_filtered, aes(x = Year, y = Current_health_expenditure_per_capita_current_US, color = CountryName, group = CountryName)) +
  geom_line(size = 1) +
  geom_point(size = 2) +  
  labs(
    title = "Current Per Capita Health Expenditure (2010–2020): Top 5 nations",
    x = "Year",
    y = "Health Expenditure (% of GDP)",
    color = "Country"
  ) +
  scale_y_continuous(breaks = y_breaks) +  
  theme_minimal()

@@

@@

```{r}
# Define dataset

@@

# Arrange plots in a 2x2 grid
final_plot1 <- (plot1 + plot2) / (plot3 + plot4)
png_file1 <- here("results", "figures", "correlation_plot1.png")
ggsave(filename = png_file1, plot = final_plot1, width = 10, height = 8, dpi = 300)

# Generate individual plots with different colors
plot5 <- create_corr_plot("Literacy_rate_adult_total__of_people_ages_15_and_above", df, "green")
plot6 <- create_corr_plot("Trade__of_GDP", df, "green")
plot7 <- create_corr_plot("Foreign_direct_investment_net_inflows_BoP_current_US", df, "green")
plot8 <- create_corr_plot("Age_dependency_ratio__of_workingage_population", df, "green")

# Arrange plots in a 2x2 grid
final_plot2 <- (plot5 + plot6) / (plot7 + plot8)
png_file2 <- here("results", "figures", "correlation_plot2.png")
ggsave(filename = png_file2, plot = final_plot2, width = 10, height = 8, dpi = 300)
```

>>>>

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(patchwork)
library(kableExtra)
library(htmltools)
library(magrittr)
library(webshot2)
library(patchwork)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
```

@@

# creating a summary table, excluding specific rows
country_summary_table <- data_2010 %>%
  filter(!CountryName %in% c("Data from database: World Development Indicators", "OECD members",        "Last Updated: 01/28/2025")) %>%
  select(
    CountryName,
    GDP_current_US,
    Population_total,
    Life_expectancy_at_birth_total_years,
    Physicians_per_1000_people,
    Hospital_beds_per_1000_people,
    Government_Effectiveness_Estimate,
    Control_of_Corruption_Estimate,
    Rule_of_Law_Estimate,
    Access_to_electricity__of_population,
    Unemployment_total__of_total_labor_force_modeled_ILO_estimate,
    Poverty_headcount_ratio_at_national_poverty_lines__of_population
  )

# generate the HTML table
table_html <- country_summary_table %>%
  kable("html", escape = FALSE, digits = 2) %>%
  kable_styling(full_width = FALSE, position = "left") %>%
  column_spec(1, bold = TRUE) %>%
  add_header_above(c(" " = 1, "National Development Indicators" = 11)) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# saving the table object as an RDS file
rds_file <- here::here("results", "tables", "health_exp_2010.rds")
saveRDS(table_html, file = rds_file)

@@

# Creating the line plot
p1<- ggplot(data_filtered, aes(x = Year, y = Current_health_expenditure__of_GDP, color = CountryName, group = CountryName)) +
  geom_line(size = 1) +
  geom_point(size = 2) +  
  labs(
    title = "Current Health Expenditure as a % of GDP (2010–2020): Top 5 nations",
    x = "Year",
    y = "Health Expenditure (% of GDP)",
    color = "Country"
  ) +
  scale_y_continuous(breaks = y_breaks) +  
  theme_minimal()

@@

The Health Expenditure Per Capita (USD) varriable in fact shows more variation and also show some connections with annual population growth.

@@

# Creating the line plot
p2<- ggplot(data_filtered, aes(x = Year, y = Current_health_expenditure_per_capita_current_US, color = CountryName, group = CountryName)) +
  geom_line(size = 1) +
  geom_point(size = 2) +  
  labs(
    title = "Current Per Capita Health Expenditure (2010–2020): Top 5 nations",
    x = "Year",
    y = "Per capita Health Expenditure (USD)",
    color = "Country"
  ) +
  scale_y_continuous(breaks = y_breaks) +  
  theme_minimal()

@@

saving the health expenditure plot

```{r}
combined_plot <- p1 / p2 + 
plot_layout(widths = c(1, 1)) +  # Equal width for both plots
plot_annotation(tag_levels = 'A')

ggsave(
  filename = here("results", "figures", "combined_health_expenditure_plots.png"),
  plot = combined_plot,
  width = 10,  #  width as needed
  height = 8,  #  height as needed
  dpi = 300
)
```

@@

```{r}
# defining dataset

@@

# arranging plots in a 2x2 grid
final_plot1 <- (plot1 + plot2) / (plot3 + plot4)
png_file1 <- here("results", "figures", "correlation_plot1.png")
ggsave(filename = png_file1, plot = final_plot1, width = 10, height = 8, dpi = 300)

# generating individual plots with different colors
plot5 <- create_corr_plot("Literacy_rate_adult_total__of_people_ages_15_and_above", df, "green")
plot6 <- create_corr_plot("Trade__of_GDP", df, "green")
plot7 <- create_corr_plot("Foreign_direct_investment_net_inflows_BoP_current_US", df, "green")
plot8 <- create_corr_plot("Age_dependency_ratio__of_workingage_population", df, "green")

# arranging plots in a 2x2 grid
final_plot2 <- (plot5 + plot6) / (plot7 + plot8)
png_file2 <- here("results", "figures", "correlation_plot2.png")
ggsave(filename = png_file2, plot = final_plot2, width = 10, height = 8, dpi = 300)
```

# EDA (part 3)

## Figures for manuscript

#cleaning and preparing the data
trend_data <- train %>%
  # Remove non-country rows
  filter(!CountryName %in% c(
    "Data from database: World Development Indicators",
    "OECD members",
    "Last Updated: 01/28/2025")) %>%

  # renaming columns
  select(
    year = Year,
    country = CountryName,
    gdp = GDP_current_US,
    population = Population_total,
    life_exp = Life_expectancy_at_birth_total_years,
    physicians = Physicians_per_1000_people,
    hospital_beds = Hospital_beds_per_1000_people,
    gov_effectiveness = Government_Effectiveness_Estimate,
    corruption = Control_of_Corruption_Estimate,
    rule_of_law = Rule_of_Law_Estimate,
    electricity = Access_to_electricity__of_population,
    unemployment = Unemployment_total__of_total_labor_force_modeled_ILO_estimate,
    poverty = Poverty_headcount_ratio_at_national_poverty_lines__of_population
  ) %>%

  #converting to long format
  pivot_longer(
    cols = -c(year, country),
    names_to = "indicator",
    values_to = "value"
  ) %>%

  #calculating yearly averages
  group_by(year, indicator) %>%
  summarise(avg_value = mean(value, na.rm = TRUE)) %>%
  ungroup()

#making a reusable plotting function
create_trend_plot <- function(data, indicator_name, plot_title, y_label, color) {
  data %>%
    filter(indicator == indicator_name) %>%
    ggplot(aes(x = year, y = avg_value)) +
    geom_line(color = color, linewidth = 1) +
    geom_point(color = color, size = 2) +
    labs(title = plot_title, y = y_label, x = NULL) +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(size = 11))
}

#now defining colors for consistency
plot_colors <- c(
  "#1f77b4", "#ff7f0e", "#2ca02c", "#d62728",
  "#9467bd", "#8c564b", "#e377c2", "#7f7f7f",
  "#bcbd22", "#17becf", "#393b79"
)

#generating the plots
plots <- list(
  gdp = create_trend_plot(trend_data, "gdp", "GDP (current US$)", "USD", plot_colors[1]),
  population = create_trend_plot(trend_data, "population", "Total Population", "Count", plot_colors[2]),
  life_exp = create_trend_plot(trend_data, "life_exp", "Life Expectancy", "Years", plot_colors[3]),
  physicians = create_trend_plot(trend_data, "physicians", "Physicians per 1000", "Count", plot_colors[4]),
  hospital_beds = create_trend_plot(trend_data, "hospital_beds", "Hospital Beds per 1000", "Count", plot_colors[5]),
  gov_effectiveness = create_trend_plot(trend_data, "gov_effectiveness", "Government Effectiveness", "Score", plot_colors[6]),
  corruption = create_trend_plot(trend_data, "corruption", "Control of Corruption", "Score", plot_colors[7]),
  rule_of_law = create_trend_plot(trend_data, "rule_of_law", "Rule of Law", "Score", plot_colors[8]),
  electricity = create_trend_plot(trend_data, "electricity", "Rural Electricity Access", " Percentage", plot_colors[9]),
  unemployment = create_trend_plot(trend_data, "unemployment", "Unemployment Rate", "% Labor Force", plot_colors[10]),
  poverty = create_trend_plot(trend_data, "poverty", "Poverty Headcount", "% of Population", plot_colors[11])
)

#combining plots ----
#create a 3-column layout 
combined_plot <- wrap_plots(plots, ncol = 3) +
  plot_annotation(
    title = "Trends in Key Development Indicators",
    subtitle = "Country averages across time",
    caption = "Source: World Development Indicators",
    theme = theme(
      plot.title = element_text(size = 16, face = "bold"),
      plot.subtitle = element_text(size = 12)
    )
  )

#saving the plot
ggsave(
  filename = here::here("results", "figures", "development_indicators_trends.png"),
  plot = combined_plot,
  width = 16,
  height = 14,
  dpi = 300
)

### 2025-04-19 code/processing-code/processingfile-v1.qmd
part #4
project #5

@@

Here, we will transform the variables as column names with appropriates values assigned to them correctly.
```{r}
# Spread the SeriesName into separate columns
data_final <- data_long %>%
  pivot_wider(
    names_from = SeriesName,  # Use SeriesName to create new columns
    values_from = Value       # Fill the new columns with the corresponding values
  )

@@

@@

@@

```{r}
xl_path <- here("data", "processed-data", "processeddata.xlsx")
writexl::write_xlsx(processeddata, xl_path)
```

>>>>

Please see the readme file before running the codes.

@@

Here, we will transform the variables as column names with appropriates values assigned to them correctly.

CAUTION: This chunck might take a long time to process. Please wait pariently for few minutes (2-5 minutes if your PC has 16gb RAM) until the code processes itself.

```{r}
# Spread the SeriesName into separate columns
data_final <- data_long %>%
  pivot_wider(
    names_from = SeriesName,  # Use SeriesName to create new columns
    values_from = Value       # Fill the new columns with the corresponding values
  )

@@

CAUTION: This chunck might also consume a good amoung of your PC's memory, expect upto 1-2 minutes for the code to process.

@@

CAUTION: This chunk could also take upto 1-2 minutes to process.

@@

```{r}
xl_path <- here("data", "processed-data", "processeddata.xlsx")
writexl::write_xlsx(data_collapsed, xl_path)
```

### 2025-04-19 code/robustness-code/robustness-code.qmd
project #5

```{r}
# Loading required packages
library(purrr)    # For map_df()
library(dplyr)    # For data manipulation
library(timetk)   # For time_series_cv()
library(rsample)  # For analysis() and assessment()
library(here)
library(modelsummary)
library(tidyr)
library(ranger)

@@

@@

# cross validation

@@

#saving the results
results <- cv(cv_splits$splits, model_formula)
print(results)
```

@@

print(cv_summary)
```

@@

```{r}
rf_cv <- function(splits, model_formula) {
  map_df(splits, function(split) {
    train_cv <- analysis(split)
    test_cv <- assessment(split)

@@

@@

```{r}
#calculating metrics
performance_metrics <- test_results %>%
  summarise(
    RMSE = sqrt(mean(Residual^2, na.rm = TRUE)),
    MAE = mean(abs(Residual), na.rm = TRUE),
    R2 = 1 - (sum(Residual^2) / sum((.data[[target_var]] - mean(.data[[target_var]]))^2))
  )

# printing metrics
cat("\nTest Set Performance:\n")
print(performance_metrics)

@@

```{r}
#viewing the predictions vs actuals
head(test_results %>% select(all_of(target_var), Predicted, Residual))
```

>>>>

NOTE: Not all codes may produce an output, you can find those outputs saved in the respective folder location.

```{r}
# Loading required packages
library(purrr)    # For map_df()
library(dplyr)    # For data manipulation
library(timetk)   # For time_series_cv()
library(rsample)  # For analysis() and assessment()
library(here)
library(modelsummary)
library(tidyr)
library(ranger) 
library(knitr)
library(kableExtra)

```
# Importing dataset

@@

# OLS regression on dataset

@@

# cross validation of OLS model

@@

#saving the results
results <- cv(cv_splits$splits, model_formula)
print(results)

#saving the results
#saving the table
cv_results <- kable(results, format = "html", digits = 3) %>%
kable_styling()
save_kable(cv_results, file = here("results", "tables","OLS-CV-results-full.html"))
```

Now we will see the summary of the CV results

@@

print(cv_summary)

#saving the table
cv_summary_kable <- kable(cv_summary, format = "html", digits = 3) %>%
kable_styling()
save_kable(cv_summary_kable, file = here("results", "tables","OLS-CV-results.html"))
```

@@

## Random Forest using train dataset

target_var <- all.vars(model_formula)[1]
train_processed <- train %>%
  drop_na(all_of(target_var))

# Train Random Forest on full training dataset
set.seed(123)
rf_full_model <- ranger(
  formula = model_formula,
  data = train_processed,
  num.trees = 500,
  mtry = mtry_value,
  importance = 'permutation',
  seed = 123
)

# Extract variable importance
importance_df <- data.frame(
  Variable = names(rf_full_model$variable.importance),
  Importance = rf_full_model$variable.importance
) %>% 
  arrange(desc(Importance))

# Create importance plot
library(ggplot2)
importance_plot <- ggplot(importance_df, aes(x = reorder(Variable, Importance), 
                         y = Importance)) +  # Fixed this line - removed extra parenthesis
  geom_col(fill = "steelblue") +  # Simplified this line
  coord_flip() +
  labs(title = "Random Forest Variable Importance",
       x = "Predictor Variables",
       y = "Importance (Permutation)") +
  theme_minimal()

# Save the plot
ggsave(here("results", "figures", "rf_importance_plot.png"), 
       plot = importance_plot,
       width = 8, height = 6, dpi = 300)
```

Here we will calcualte the performance metrics of our RF model.

```{r}
train_predictions <- predict(rf_full_model, data = train_processed)$predictions

# Calculate performance metrics
train_performance <- train_processed %>%
  mutate(Predicted = train_predictions) %>%
  summarise(
    RMSE = sqrt(mean((.data[[target_var]] - Predicted)^2)),
    MAE = mean(abs(.data[[target_var]] - Predicted)),
    R2 = cor(.data[[target_var]], Predicted)^2
  ) %>%
  mutate(across(everything(), ~round(., 3)))  # Round all metrics to 3 decimal places

# Display the performance metrics
print(train_performance)

Saving the table

```{r}
library(kableExtra)
 train_table <- train_performance %>%
  pivot_longer(everything(), names_to = "Metric", values_to = "Value") %>%
  mutate(Metric = case_when(
    Metric == "RMSE" ~ "Root Mean Squared Error",
    Metric == "MAE" ~ "Mean Absolute Error",
    Metric == "R2" ~ "R-squared",
    TRUE ~ Metric
  )) %>%
  kable(format = "html", align = c("l", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

kableExtra::save_kable(train_table, 
                       file = here("results", "tables", "rf_train_performance.html"))
```

## Random Forest with CV on train dataset

```{r}
rf_cv <- function(splits, model_formula, include_importance = TRUE) {
  map_df(splits, function(split) {
    train_cv <- analysis(split)
    test_cv <- assessment(split)

@@

Saving the RF results

```{r}
# Select only the desired columns
rf_results_table <- rf_results %>%
  select(Train_Years, Test_Year, RMSE, Predictors_Used, mtry_Used)

# Create and save the HTML table
rf_results_html <- kable(rf_results_table, 
                         format = "html",
                         digits = 2,
                         caption = "") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "left") %>%
  add_header_above(c("Training Period" = 1, "Test Year" = 1, 
                     "Performance" = 1, "Model Parameters" = 2))

# Save the table
save_kable(rf_results_html, 
           file = here("results", "tables", "rf_cv_results.html"))
```

Now we will calculate the summary of the above results.

# Create summary table first
rf_summary_table <- rf_results %>%
  summarise(
    `Mean RMSE` = mean(RMSE, na.rm = TRUE),
    `SD RMSE` = sd(RMSE, na.rm = TRUE),
    `Min RMSE` = min(RMSE, na.rm = TRUE),
    `Max RMSE` = max(RMSE, na.rm = TRUE)
  ) %>%
  mutate(across(everything(), ~round(., 3)))  # Round all values

# Then create and save HTML table
rf_summary_html <- kable(rf_summary_table, 
                         format = "html",
                         caption = "") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "center") %>%
  row_spec(0, bold = TRUE)

save_kable(rf_summary_html, file = here("results", "tables", "rf_cv_summary.html"))
```

@@

```{r}
# Calculate metrics manually
performance_metrics <- data.frame(
  RMSE = sqrt(mean(test_results$Residual^2)),
  MAE = mean(abs(test_results$Residual)),
  R2 = cor(test_results[[target_var]], test_results$Predicted)^2
)

print(performance_metrics)
```

# Format the performance metrics into a clean table
performance_table <- performance_metrics %>%
  pivot_longer(cols = everything(), 
               names_to = "Metric", 
               values_to = "Value") %>%
  mutate(Metric = case_when(
    Metric == "RMSE" ~ "Root Mean Squared Error",
    Metric == "MAE" ~ "Mean Absolute Error",
    Metric == "R2" ~ "R-squared",
    TRUE ~ Metric
  ),
  Value = round(Value, 3))  # Round to 3 decimal places

# Create and save the HTML table
performance_html <- kable(performance_table, 
                         format = "html",
                         col.names = c("Performance Metric", "Value"),
                         align = c("l", "r"),
                         caption = "") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "center") %>%
  row_spec(0, bold = TRUE, background = "#f8f9fa") %>%
  column_spec(2, width = "5cm")

# Save the table
save_kable(performance_html, 
           file = here("results", "tables", "rf_test_performance.html"))

@@

# Visualizing the models train dataset

```{r}
# For Random Forest model 
rf_train_predictions <- predict(rf_full_model, data = train_processed)$predictions

# For OLS model 
ols_model <- lm(model_formula, data = train_processed)
ols_train_predictions <- predict(ols_model, newdata = train_processed)

#creating combined data frame for plotting
plot_data <- train_processed %>%
  mutate(
    RF_Predicted = rf_train_predictions,
    OLS_Predicted = ols_train_predictions,
    Observed = .data[[target_var]]
  )

#creating the comparison plot
library(ggplot2)
pred_vs_obs_plot <- ggplot(plot_data, aes(x = Observed)) +
  geom_point(aes(y = RF_Predicted, color = "Random Forest"), alpha = 0.6) +
  geom_point(aes(y = OLS_Predicted, color = "OLS Regression"), alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  labs(
    title = "Predicted vs Observed Values (Training Data)",
    x = "Observed Health Expenditure per Capita (US$)",
    y = "Predicted Health Expenditure per Capita (US$)",
    color = "Model Type"
  ) +
  scale_color_manual(values = c("Random Forest" = "#1f77b4", 
                               "OLS Regression" = "#ff7f0e")) +
  theme_minimal() +
  theme(legend.position = "bottom")

#saing the plot
ggsave(here("results", "figures", "pred_vs_obs_train.png"),
       plot = pred_vs_obs_plot,
       width = 8, height = 6, dpi = 300)

# Display  plot
print(pred_vs_obs_plot)
```

# Visualizing the models test dataset

```{r}
#generating predictions for test data as we did for train data

# Random Forest predictions 
rf_test_predictions <- predict(rf_model, data = test_processed)$predictions

# OLS predictions on test data
ols_test_predictions <- predict(model, newdata = test_processed)  #

# combined data frame for plotting
test_plot_data <- test_processed %>%
  mutate(
    RF_Predicted = rf_test_predictions,
    OLS_Predicted = ols_test_predictions,
    Observed = .data[[target_var]]
  )

# comparison plot
test_pred_vs_obs_plot <- ggplot(test_plot_data, aes(x = Observed)) +
  geom_point(aes(y = RF_Predicted, color = "Random Forest"), alpha = 0.6, size = 2) +
  geom_point(aes(y = OLS_Predicted, color = "OLS Regression"), alpha = 0.6, size = 2) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  labs(
    title = "Predicted vs Observed Values (Test Data)",
    x = "Observed Health Expenditure per Capita (US$)",
    y = "Predicted Health Expenditure per Capita (US$)",
    color = "Model Type"
  ) +
  scale_color_manual(values = c("Random Forest" = "#E69F00",  
                               "OLS Regression" = "#56B4E9")) +  
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5)) +  # Center title
  coord_equal()  # Equal axis scaling for proper 45° line

#adding performance metrics to plot
test_performance <- test_plot_data %>%
  summarise(
    RF_RMSE = sqrt(mean((Observed - RF_Predicted)^2)),
    OLS_RMSE = sqrt(mean((Observed - OLS_Predicted)^2))
  )

test_pred_vs_obs_plot <- test_pred_vs_obs_plot +
  annotate("text", x = min(test_plot_data$Observed), 
           y = max(c(test_plot_data$RF_Predicted, test_plot_data$OLS_Predicted)),
           hjust = 0, vjust = 1,
           label = sprintf("RMSE:\nRandom Forest: %.1f\nOLS: %.1f",
                          test_performance$RF_RMSE, 
                          test_performance$OLS_RMSE))

#saving the plot
ggsave(here("results", "figures", "pred_vs_obs_test.png"),
       plot = test_pred_vs_obs_plot,
       width = 8, height = 6, dpi = 300)

#show the plot
print(test_pred_vs_obs_plot)
```
